### 2015-09-05 practice/docs/thesis/nomenclature.tex
hopefully this is enough to compile final thesis draft?

\clearpage
\section*{\huge{Nomenclature}}

\begin{longtable}{ p{.15\textwidth}  p{.85\textwidth} }
$a$							& 	Two-form in model weak form\\
$\mathcal{C}$		&		Observation operator \\
$d$							&		Observations \\
$e$							&		Error in stationary point of augmented Lagrangian \\
$f$							&		Forcing function \\
$HF$						&		High-fidelity \\
$I$							&		QoI functional \\
$J$							&		Objective function of optimization formulation of inverse problem \\
$k_d$						&		Diffusion coefficient \\
$k_r$						&		Reaction coefficient \\
$\ell$					&		Functional in model weak form \\
$\mathcal{L}$		&		Lagrangian \\
$LF$						&		Low-fidelity \\
$\mathcal{M}$		&		Augmented Lagrangian \\
$MF$						&		Mixed-fidelity \\
$n_d$						&		Number of observations \\
$p$							&		Auxiliary variable corresponding to parameters \\
$q$   					&   Unknown parameters \\
$Q$							&		Hilbert space \\
$\mathcal{Q}$		&		Output functional \\
$R$							&		Regularization term \\
$\mathcal{R}$		&		Higher-order remainder terms \\
$\mathscr{R}$		&		Residual form for QoI error adjoint \\
$u$   					&   State variables \\
$U$							&		Hilbert space \\
$v$							&		Auxiliary variable corresponding to state \\
$\vec{V}$				&		Velocity field \\
$x_1,\:x_2$			&		Spatial coordinates \\
$y$							&	 	Auxiliary variable corresponding to adjoint \\
$z$	  					&	  Adjoint variables \\
$\beta$					&		Regularization coefficient \\
$\Lambda$				&		Supplementary adjoint \\
$\xi$						&		Primary variables \\
$\phi$					&		Test function \\
$\Phi$					&		Test function \\
$\chi$					&		Auxiliary variables \\
$\Psi$					&		Stationary point of augmented Lagrangian \\
$\Omega$				&		Domain \\
$\Omega_{HF}$		&		Subdomain over which high-fidelity model is solved \\
$\Omega_I$			&		Region in domain over which QoI is calculated \\
$\Omega_{LF}$		&		Subdomain over which low-fidelity model is solved \\
 & \\
$a'_u(u_0)(\delta u)$	& Fr\'{e}chet derivative of form $a$ with respect to $u$, at $u_0$, in the direction $\delta u$
\end{longtable}