### 2016-05-21 paper/abstract.tex
Reorged math formulation to make more concise for journal, express main result in theorem form Added a new algorithms section with main algorithm and complexity analysis Added abstract

Inference problems are often constrained by state equations, which arise from conservation laws, and constitutive relations. Often, a hierarchy of models can be derived from such laws and relations, reflecting varying fidelity versus cost tradeoffs. We introduce a goal oriented, model adaptive, inference algorithm, which allows users to achieve an aritrary error tolerance, while minmizing the use of high fidelity constraints. Numerical experiments test the algorithm on a highly nonlinear inverse problem, and showcase the computational savings and robustness benefits of this new approach.