### 2016-07-17 paper/conclusions.tex
edits to present results in slightly different light

\section{Conclusion}\label{sect:conc}
Future work:
\begin{itemize}
\item stochastic?
\item mixing models in time as well? (different mixes of models at different time steps?)
\end{itemize}