### 2016-09-29 paper/abstract.tex
small paper edits

\red{Inference problems are often constrained by state equations, which arise from conservation laws, and constitutive relations. Often, a hierarchy of models can be derived from such laws and relations, reflecting varying fidelity versus cost tradeoffs. We introduce a goal-oriented, model adaptive, inference algorithm, which allows users to achieve an arbitrary error tolerance, while minimizing the use of high fidelity constraints. Numerical experiments exercise the algorithm on a highly nonlinear inverse problem, and showcase the computational savings and robustness benefits of this new approach.}