### 2016-09-30 paper/abstract_v1.tex
renamed paper tex files so that we can do latexdiff on them

Inference problems are often constrained by state equations, which arise from conservation laws, and constitutive relations. Often, a hierarchy of models can be derived from such laws and relations, reflecting varying fidelity versus cost tradeoffs. We introduce a goal oriented, model adaptive, inference algorithm, which allows users to achieve an arbitrary error tolerance, while minimizing the use of high fidelity constraints. Numerical experiments exercise the algorithm on a highly nonlinear inverse problem, and showcase the computational savings and robustness benefits of this new approach.