### 2016-11-10 paper/paper_v3.tex

\documentclass[review]{siamart0516}
%%\usepackage{lgrind}

%\usepackage[square,comma,numbers,sort&compress]{natbib}
%\usepackage{amsmath,amsthm,amssymb,amsfonts,amsbsy,latexsym} %amsthm inclusion conflicts with sisc...
\usepackage{amsmath,amssymb,amsfonts,amsbsy,latexsym}
\usepackage{graphics}
%\usepackage{epsfig}
%%\usepackage[hang,raggedright]{subfigure}
%\usepackage{epsf}
%\usepackage{setspace}
%\usepackage{hangcaption}
\usepackage{graphicx}    % needed for including graphics e.g. EPS, PS
\usepackage{multirow}
%\usepackage{threeparttable}
\usepackage{cancel}
\usepackage{enumerate}
\usepackage{color}
\usepackage{hyperref}

%\usepackage{algorithmicx}
%\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
\usepackage{varwidth}
%\usepackage{caption}
%\usepackage[font=small]{subcaption}
%\usepackage{longtable}
\usepackage{mathrsfs}
\usepackage{upgreek}
\usepackage[caption=false]{subfig} % for \subfloat

\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links
    citecolor=black,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=blue           % color of external links
}

\newcommand{\R}{{\mathbb{R}}}
\newcommand{\Reals}{{\mathbb{R}}}
\newcommand{\bit}{\begin{itemize}}
\newcommand{\eit}{\end{itemize}}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\orderof}[1]{{\ensuremath{ {\cal O}(#1)}}}

%\newtheorem{theorem}{Theorem} %conflicts with sisc

\graphicspath{{./figures/}}

\begin{document}
\raggedbottom %avoid weird vertical justification

\title{Model Adaptivity for Goal-Oriented Inference using Adjoints}

\author{Harriet Li, Vikram Garg, Karen Willcox}
%\date{}
\maketitle

\begin{abstract}
\input{abstract_v2.tex}
\end{abstract}

\begin{keywords}
  inference, goal-oriented adaptive modeling, a posteriori error estimation, multifidelity modeling, adjoints, PDEs
\end{keywords}

\begin{AMS}
  35R30, 65G20, 62F35, 35R15
\end{AMS}

\input{intro_v2.tex}
\input{mathematical_formulation_v2.tex}
\input{algorithms_v2.tex}
\input{numerical_experiments_v2.tex}
\input{conclusions_v2.tex}
\input{acknow.tex}

\bibliographystyle{siamplain}
\bibliography{masterBib}
\end{document}