### 2016-12-14 paper/paper_v4_cmame.tex
fixing latex formatting issues in cmame version

\documentclass[review,sort&compress]{elsarticle}
%%\usepackage{lgrind}

%\usepackage[square,comma,numbers,sort&compress]{natbib}
\usepackage{amsmath,amsthm,amssymb,amsfonts,amsbsy,latexsym} %amsthm inclusion conflicts with sisc...
%\usepackage{amsmath,amssymb,amsfonts,amsbsy,latexsym}
\usepackage{graphics}
%\usepackage{epsfig}
%%\usepackage[hang,raggedright]{subfigure}
%\usepackage{epsf}
%\usepackage{setspace}
%\usepackage{hangcaption}
\usepackage{graphicx}    % needed for including graphics e.g. EPS, PS
\usepackage{multirow}
%\usepackage{threeparttable}
\usepackage{cancel}
\usepackage{enumerate}
\usepackage{color}
\usepackage{hyperref}
\usepackage{cleveref}

\usepackage{algorithmicx}
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
\usepackage{varwidth}
%\usepackage{caption}
%\usepackage[font=small]{subcaption}
%\usepackage{longtable}
\usepackage{mathrsfs}
\usepackage{upgreek}
\usepackage[caption=false]{subfig} % for \subfloat

\hypersetup{
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=black,          % color of internal links
    citecolor=black,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=blue           % color of external links
}

\newcommand{\R}{{\mathbb{R}}}
\newcommand{\Reals}{{\mathbb{R}}}
\newcommand{\bit}{\begin{itemize}}
\newcommand{\eit}{\end{itemize}}
\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\orderof}[1]{{\ensuremath{ {\cal O}(#1)}}}

\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}

\graphicspath{{./figures/}}

\begin{document}
%\raggedbottom %avoid weird vertical justification

\begin{frontmatter}

\title{Model Adaptivity for Goal-Oriented Inference using Adjoints}

%\author{Harriet Li\textsuperscript{a,}\footnote{Corresponding author. \textit{Email address: kameeko@mit.edu}}, Vikram Garg\textsuperscript{b}, Karen Willcox\textsuperscript{c}}
%\address{\textsuperscript{a} Department of Aeronautics and Astronautics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA; \texttt{kameeko@mit.edu}\protect\\
%\textsuperscript{b} Institute for Computational Engineering and Sciences, University of Texas, Austin, TX, 78712, USA; \texttt{vikram.v.garg@gmail.com}\protect\\
%\textsuperscript{c} Department of Aeronautics and Astronautics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA; \texttt{kwillcox@mit.edu}}
%\date{}
%\maketitle

\author[adr1]{Harriet Li\corref{cor1}}
\cortext[cor1]{Corresponding author. \textit{Email address: kameeko@mit.edu}}
\address[adr1]{Department of Aeronautics and Astronautics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA; \texttt{kameeko@mit.edu}}
\author[adr2]{Vikram Garg}
\address[adr2]{Institute for Computational Engineering and Sciences, University of Texas, Austin, TX, 78712, USA; \texttt{vikram.v.garg@gmail.com}}
\author[adr3]{Karen Willcox}
\address[adr3]{Department of Aeronautics and Astronautics, Massachusetts Institute of Technology, Cambridge, MA, 02139, USA; \texttt{kwillcox@mit.edu}}

\begin{abstract}
\input{abstract_v4.tex}
\end{abstract}

\begin{keyword}
  inference \sep goal-oriented adaptive modeling \sep a posteriori error estimation \sep multi-fidelity modeling \sep adjoints
\end{keyword}

\end{frontmatter}

\input{intro_v4_cmame.tex}
\input{mathematical_formulation_v4_cmame.tex}
\input{algorithms_v4_cmame.tex}
\input{numerical_experiments_v4_cmame.tex}
\input{conclusions_v4_cmame.tex}
\input{acknow.tex}

\section*{References}
%\bibliographystyle{plain}
\bibliographystyle{elsarticle-harv} %this seems to get initials in bibliography, without doing name-year citation style as bst file claims?
\bibliography{masterBib}
\end{document}